---
title: "Final Project"
author: Lauryn Pavlik
output: html_notebook
---
#Make sure to explain data sets and variables




#Clean environment and load packages
```{r}
#Clean up workspace environment
rm(list = ls())

#All packages used for the project
library(mosaic)
library(tidyr)
library(DataComputing)
library(readr)
library(dplyr)
library(ggplot2)


```

#Import first data set and name it 
```{r}
library(readr)
new_york <- read_csv("final decision.csv")
```

#View and make data table tidy 
```{r}
ny_data <- 
new_york %>%
 dplyr:: filter(Grade!="All Grades") %>%
  rename(district= "District", grade= "Grade", year= "Year", mean = "Mean Scale Score", level_1 = "Level 1 #", level_1_perc = "Level 1 %", level_2 = "Level 2 #", level_2_perc = "Level 2 %",level_3 = "Level 3 #", level_3_perc = "Level 3 %",level_4 = "Level 4 #", level_4_perc = "Level 4 %", students="Number Tested", sex= "Category" )
  
ny_data$`Level 3+4 #` <- NULL
ny_data$`Level 3+4 %` <- NULL




```

#Play around with data and find any correlation
##Does the number of students taking the benchmark increase the average score? 
```{r}
ny_data %>%
  ggplot(aes(x= students, y= mean)) + 
  geom_point(aes(color=sex))



```
##Do females tend to have a higher score than males?
```{r}
ny_data %>%
  ggplot(aes(x= mean, y= district)) + 
  geom_boxplot() +
  facet_grid(year ~ sex)



```





#Hypothesis test to see if the majority of students are a level 3
```{r}


```


#Introduce Conneticut data set 



#Play around with it 


#Regression model


#Tree prediction 

#Build a function? 
















