---
title: "Final Project"
author: Lauryn Pavlik
output: html_notebook
---
#Make sure to explain data sets and variables




#Clean environment and load packages
```{r}
#Clean up workspace environment
rm(list = ls())

#All packages used for the project
library(mosaic)
library(tidyr)
library(DataComputing)
library(readr)
library(dplyr)
library(ggplot2)


```

#Import first data set and name it 
```{r}
library(readr)
new_york <- read_csv("final decision.csv")
```

#View and make data table tidy 
```{r}
ny_data <- 
new_york %>%
 dplyr:: filter(Grade!="All Grades") %>%
  rename(district= "District", grade= "Grade", year= "Year", mean = "Mean Scale Score", level_1 = "Level 1 #", level_1_perc = "Level 1 %", level_2 = "Level 2 #", level_2_perc = "Level 2 %",level_3 = "Level 3 #", level_3_perc = "Level 3 %",level_4 = "Level 4 #", level_4_perc = "Level 4 %", students="Number Tested", sex= "Category" )
  
ny_data$`Level 3+4 #` <- NULL
ny_data$`Level 3+4 %` <- NULL




```

#Play around with data and find any correlation
##Does the number of students taking the benchmark increase the average score? 
```{r}
ny_data %>%
  ggplot(aes(x= students, y= mean)) + 
  geom_point(aes(color=sex))



```

There is a slight postive correlation, but nothing too signficant.
##Let's add more variables and see if there is any correlation between more and new variables.  
```{r}
ny_data %>%
  ggplot(aes(x= district, y= mean)) + 
  geom_point() +
  facet_grid(sex~year)




```
#Judging from the past graph, it seems that females have a slightly higher benchmark score. It is also noticeable that the scores over the districts in the more recent years have less of a postive correlation.These observations bring about two questions: 

##1.Using a hypothesis test, is there statistically significant evidence that females have higher scores than males? 
```{r}
m_s<- 
  ny_data[,1:6]

ms_wide <- 
  m_s %>%
data.frame() %>%
  spread(key= sex, value=mean)

mu1 <- ms_wide[[5, na.rm=TRUE]]
mu2 <- c(ms_wide$Male, na.rm = TRUE)
x_bar1 <- mean(ms_wide$Female, na.rm = TRUE)
x_bar2 <- mean(ms_wide$Male, na.rm = TRUE)
x_bar <- x_bar1 - x_bar2
dif_of_mean <- c(ms_wide$Female, na.rm = TRUE)






```


```{r}



```

##2. Based on a regression model, what variables influence mean score the most?
```{r}





```

















