---
title: "Final Project"
author: "Lauryn Pavlik"
output: html_notebook
---
# Background of The Data
The data set that is being used in this analysis is taken from data.gov. This set evolved from math benchmark data in New York. It contains information such as average score, number of participants, how many students scored in each category, and the percent of students in each category. Level one is the lowest score range and level four is the highest. This data is collected from 32 districts and six grade levels over the course of six years. 

## Clean Environment and Load Packages
```{r message = FALSE}

# Clean up workspace environment
rm(list = ls())

# All packages used for the project
library(tidyr)
library(dplyr)
library(ggplot2)
library(rpart.plot)
library(rpart)
library(readr)

```

## Import Data  
```{r message = FALSE}
# Get data
New_york <- read_csv("final decision.csv")

# View data table
head(New_york)

```

Here the table is not yet tidy because all variable names to need to be editied to fit the code style. 

## Inspect Bottom of Data Table 
```{r}
# View last six rows
tail(New_york)

```

By viewing the last six rows of the data table, one can see that the districts extend all the way to 32 and the year goes to 2011. Although, the data for the Grade variable does not contain a single grade level for each case. This is one of the many errors that will need tidying. 

## Fix Data Table 
```{r warning = FALSE}

# Rename the variables to accommodate style conventions
Ny_data <- 
New_york %>%
 dplyr:: filter(Grade != "All Grades") %>%
  rename(district = "District", grade = "Grade", year = "Year", avg = "Mean Scale Score", level_1 = "Level 1 #", level_1_perc = "Level 1 %", level_2 = "Level 2 #", level_2_perc = "Level 2 %", level_3 = "Level 3 #", level_3_perc = "Level 3 %", level_4 = "Level 4 #", level_4_perc = "Level 4 %", students = "Number Tested", sex = "Category") 

# Remove repeated columns and non numeric inputs for quantitative variables
Ny_data$`Level 3+4 #` <- NULL
Ny_data$`Level 3+4 %` <- NULL

Ny_data %>%
  remove(row = "All Grades")

# View data table
head(Ny_data)

```

## Ask Questions To Explore Data

### What is the highest average score?
```{r}
# Create new data table with descending average score 
High <- 
Ny_data %>%
  arrange(desc(avg))
# View new data table 
head(High)

```
It is observed that the highest score is 724 points.

### What is the lowest average score?
```{r}
# See the last six rows of the descending table 

tail(High)

```
It is observed that the lowest score is 614 points.

### Does the number of students taking the benchmark increase the average score? 
```{r}
# Scatterplot 
Ny_data %>%
  ggplot(aes(x = students, y = avg)) + 
  geom_point(aes(color = sex)) +
  xlab("Number of Students") +
  ylab("Average Points Scored") +
  ggtitle("Average Score Overall")



```

It can be observed that there is a slight postive correlation, but nothing signficant to the eye. 

### Add More Variables   
```{r}
# Scatterplot
Ny_data %>% 
  ggplot(aes(x = district, y = avg)) + 
  geom_point() +
  facet_grid(sex ~ year) +   # Seperate data by year and by sex
  geom_smooth(method = lm, color = "orange") +  # Add an orange smoother to see correlation
  xlab("District Number") +
  ylab("Average Points Scored") +
  ggtitle("Gender's Average Score Over The Years")

```


Judging from the graph, it seems that females have a slightly higher benchmark score. It is also noticeable that the scores over the districts in the more recent years have less of a postive correlation. These observations bring about a question: 

## Using a hypothesis test, is there statistically significant evidence that females have higher scores than males? 
Before we can move on with the test, it is safe to check and make sure there is the same total amount of males and females that have taken the benchmark. If the results are different, the test will not be reliable. 

### Test counts
```{r}
Ny_data %>%
  group_by(sex) %>%
  summarise(count = n())


```

There is the same amount of each sex, so the test is possible. 

### First, create a data table of only female scores
```{r}
# Create table that only contains the first six variables of the original tidy table
No_levels <- 
  Ny_data[, 1:6] 

  
# Create table with only female scores
F_score <- 
  No_levels %>%
data.frame() %>%
  spread(key = sex, value = avg) %>%   # Spread sex into seperate columns
  rename(female = "Female") %>% 
  drop_na(female)   # Remove any missing values for females 

# Drop the four unneeded variables 
F_score$district <- NULL
F_score$grade <- NULL
F_score$students <- NULL
F_score$Male <- NULL




```
  
### Second, create a data table of only male scores   
```{r}
# Create table with only male scores
M_score <- 
  No_levels %>%
data.frame() %>%
  spread(key = sex, value = avg) %>%  # Again, spread sex into seperate columns
  rename(male = "Male") %>%
  drop_na(male) # Remove missing values for males

# Drop the four unneeded variables 
M_score$district <- NULL
M_score$grade <- NULL
M_score$students <- NULL
M_score$Female <- NULL


```

### Third, assign vectors to mu 1 and mu 2 and run a two sample t test
```{r}
# Assign vectors
mu_1 <- as.vector(F_score$female)
mu_2 <- as.vector(M_score$male)

# Run test
t.test(mu_1, mu_2)


```

From this hypothesis test, a p-value of 0.0003039 is produced. Since the p-value is very small, the null hypothesis can be rejected and there is significant evidence produced that females tend to have a greater average score on math benchmarks than males do. From this data, we are 95% confident that the average female score is about 1.22 points to 4.12 points higher than the male average score. 


## Now that we know females score higher than males, it is time to question the levels of scoring

### Does any level of scoring have a more significant impact than the rest on the average score? 
```{r}
# Create regression model
Avg_mod <-
  Ny_data %>%
  lm(students ~ level_1 + level_2 + level_3 + level_4, data = .)  

# View model
summary(Avg_mod)

```

The model's p-values provide evidence that all four levels of scoring contribute to the average score, as common sense suggests. When interpreting the model, the residuals are all very small values, meaning that there is little error in the calculation. This calculation shows that all levels influence the average score by the same positive amount. Regression models are great for more variant data, but are not good for closely knitted results like this as shown by the example model. 

### Take easy, but less informative route to see the influence of each level
```{r}
mean(Ny_data$level_1_perc)
mean(Ny_data$level_2_perc)
mean(Ny_data$level_3_perc)
mean(Ny_data$level_4_perc)

```

By taking the mean of the percents of the different levels, it is shown that 9.54% of students score in level one, 26.37% score in level two, 44.07% score in level three and 20.02% score in level four. It can be said that most students score in the level three zone. 

## Predict average score based on the variables
Say one knows the year, grade of the students and the percent of scores coming from each level, but needs to know the actual average score for a group of students. A regression tree is a good way to predict the average score. 

```{r}
# Create regression tree
Avg_tree <- 
Ny_data %>% 
  rpart(avg ~ . , data = .)

# Plot regression tree
rpart.plot(Avg_tree, roundint = FALSE)


```

The tree produces a model to help predict the average math benchmark score so, this would be helpful when not knowing a score for a group of students. For example, if the percent of level four scores is less than 6 and the year is 2010, we can predict the average score will be 664.


